```{r include=FALSE, cache=FALSE}
set.seed(858)

options(
  digits = 3,
  dplyr.print_max = 6,
  dplyr.print_min = 6
)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = 'center',
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)

image_dpi <- 125
image_dpi_small <- 325
```

# Organization

In this chapter, we'll talk about how to organize your project. We'll discuss how to use the project template created by `dcl::dcl_data_project()`, explain an easier way to handle file paths, and walk you through the creation of a makefile. 

## Directories

Here's an overview of the relationships between the various directories created by `dcl::dcl_data_project()`:

```{r echo=FALSE}
knitr::include_graphics(
  "images/project-workflow/folder-organization.png", dpi = image_dpi
)
```

To help explain the different directories, we'll refer to our [project example](https://github.com/dcl-docs/project-example), which uses [data](https://datadryad.org/stash/dataset/doi:10.5061/dryad.8rr0498) on migratory birds' collisions with lighted buildings in Chicago. 

### data-raw

Put all raw data files (i.e., the data that you haven't touched yet) in __data-raw__.  

For each data file, come up with a short, but descriptive, name. You'll use these names to name other files. For example, our [example data-raw](https://github.com/dcl-docs/project-example/tree/master/data-raw) contains the following raw data files:

* __birds.txt__
* __collisions.csv__
* __light_mp.csv__

### scripts

Data-manipulation scripts belong in __scripts__. We'll talk more about R scripts in [Chapter 7](https://dcl-workflow.stanford.edu/scripts.html), but you can also take a look at our [example scripts](https://github.com/dcl-docs/project-example/tree/master/scripts). 

Create one script for each raw data source. Each script should read in a raw data file, process it, and then write the cleaned data to a .rds file in __data__. 

Name each script to correspond to the data file it creates. For example, the script __collisions.R__ reads in __collisions.csv__, cleans the data, then writes _collisions.rds_ to __data__ using `readr::write_rds()`. 

If you want to join multiple data sources, create an additional script that joins the cleaned data files. For example, __bird_collisions_light.R__ joins __birds.rds__, __collisions.rds__, and __light_mp.rds__, and writes to __bird_collisions_light.rds__. As the diagram points out, your joining scripts can join cleaned data from __data__. 
  
### data

__data__ contains cleaned data (likely in .rds format) that is ready to analyze. The name of each .rds file should correspond to the script that created it. For example:

* __collisions.rds__
* __birds.rds__
* __light_mp.rds__
* __bird_collisions_light.rds__

### docs

The __docs__ folder is for any documentation files you used to understand the data, as well as for any notes you have on the data or your plan for analysis. 

### eda

The __eda__ folder contains R Markdown files with your EDA work. As a start, perform EDA on each dataset and name the R Markdown files accordingly. For example, __birds.Rmd__ performs EDA on just __birds.rds__, __bird_collisions_light.Rmd__ performs EDA on just __bird_collisions_light.rds__, etc. Once you understand your data, you can perform further EDA studies, each in an R Markdown file with a descriptive name. This separation is intended to help you organize your EDA, which can become unwieldy if placed in a single file.   

### reports

__reports__ is for your final reports, share-able analysis, etc. The files here don't need to be named according to the convention of the other files. For example, our __reports__ folder just has one report called __report.Rmd__. 

## File paths

The [here package](https://here.r-lib.org/) simplifies the work of specifying file paths. The function `here::here()` allows you to specify a file path relative to the directory of your .Rproj file, no matter what folder you're in. For example, no matter where you are in your project, you can give the file path of __collisions.rds__ as `here::here("data/collisions.rds")`. 

See the our [scripts](https://github.com/dcl-docs/project-example/tree/master/scripts), [EDA documents](https://github.com/dcl-docs/project-example/tree/master/eda), and [reports](https://github.com/dcl-docs/project-example/tree/master/reports) for more examples.

## R Markdown template

The dcl package also contains a R Markdown template to use for your EDA files and reports. To use the template:

* Click on the new file button in the top-left corner of RStudio.
* Select _R Markdown_ > _From Template_ > _DCL GitHub Document_. 

Our template is similar to the default GitHub document template, but:

* Includes a table of contents by default.
* Formats the first R chunk to highlight places for libraries, parameters, and reading in code.
* Has example headers.

## Make

Imagine that __birds.txt__, our example raw data set, gets updated. Maybe the original owners added new birds or corrected a mistake. The cleaned birds data, __birds.rds__, depends on __birds.txt__, as do other data files, EDA documents, and reports. 

To update all these files, we could rerun all our scripts, then re-knit all our EDA files and reports. However, manually updating all our files can get tedious. It also requires remembering which files depend on each other, which can get complicated. 

_Makefiles_ are a better way to update files and keep track of dependencies. A program called Make reads your makefile, looks for changes in the files specifies, then rebuilds all files that depend on the ones that changed. 

[GNU Make](https://www.gnu.org/software/make/) is free software and comes installed on Macs and most Unix machines. If you're a Windows user, you might need to install Make yourself. 

### Create a makefile

`dcl::create_data_project()` creates a makefile template for you, but you'll still need to fill in the details. We'll use our [makefile](https://github.com/dcl-docs/project-example/blob/master/Makefile) from the birds project as an example. 

Note that this section will be most informative if you have our makefile template in front of you and are ready to fill it in. 

If you want to learn more about makefiles, the [GNU Make manual](https://www.gnu.org/software/make/manual/) is a good reference.

#### Search path

Find the line that says
  
`VPATH = `

Add in the names of all the folders where Make should look for your files. This will probably be:

`VPATH = data data-raw eda reports scripts`

#### Targets

`all : $(DATA) $(EDA) $(REPORTS)` on line 14 defines a _target_ called `all`. This tells Make to, by default, consider all the files defined by `DATA`, `EDA`, and `REPORTS`. 

For this to work, you'll need to fill in the variables `DATA`, `EDA`, and `REPORTS` on lines 5-11 with the names of your files. For example, ours has

```
# Processed data files
DATA = birds.rds collisions.rds light_mp.rds bird_collisions_light.rds

# EDA studies
EDA = birds.md collisions.md light_mp.md bird_collisions_light.md

# Reports
REPORTS = report.md
```

#### Dependencies

Now, you need to specify the _dependencies_ of your project. File A depends on File B if changing File B can change File A. For example, __birds.md__ depends on __birds.rds__ because changing the cleaned data in __birds.rds__ could change the analysis, visualizations, etc. in __birds.md__.

Underneath the comment that says, `# Data dependencies`, you'll specify which how the cleaned data (the .rds files) depend on the scripts and raw data. For example, here's what we have in our birds makefile:

```
birds.rds : birds.R birds.txt
collisions.rds : collisions.R collisions.csv
light_mp.rds : light_mp.R light_mp.csv
bird_collisions_light.rds : bird_collisions_light.R \
  birds.rds collisions.rds light_mp.rds
```

The syntax is:

`[target file] : [dependency file 1] [dependency file 2] [dependency file 3]`

Next, underneath the comment that says `# EDA study and report dependencies`, specify how your knitted reports (your .md files) depend on the cleaned data.

```
birds.md : birds.rds
collisions.md : collisions.rds
light_mp.md : light_mp.rds
bird_collisions_light.md : bird_collisions_light.rds
report.md : bird_collisions_light.rds
```

Your files can have any number of dependencies, but make sure to separate the dependencies with a single space. If you need more than one line for your dependencies, end all lines except the last with a "\\".

#### Rules

Finally, Make needs to know how to update different types of files. We want Make to run a script if raw data changes, but knit an R Markdown document if cleaned data changes. Lines 30-33 define our rules. 

You probably won't need to update these rules, but it's useful to understand them. 

The first rule (lines 30-31) tells Make how to update a .rds file. For example, say __birds.txt__ changes. Make knows that __birds.rds__ depends on __birds.txt__ because of our specified dependencies. Make then looks to our first rule to figure out how to update __birds.rds__. The rule says to run the R script with the same name as the .rds file. In our example, that script is __birds.R__, so Make will run __birds.R__.

The second rule (32-33) tells Make how to update a .md file. The rule tells Make to knit the .Rmd version of the relevant .md file. For example, if __birds.md__ needs updating (because __birds.rds__ changed), Make will knit __birds.Rmd__.

### Run Make

Once you've set up your makefile (and have the necessary files in place), navigate to your project directory from the command line. Then, type `make` and hit enter. Your scripts should run and your files should knit. After `make` is done, try running `make` again. You'll get a message saying

`make: Nothing to be done for 'all'.`

because nothing has changed since the last time Make ran.

Now, you can run `make` from the command line every time you want to rebuild your project, suspect something has changed, or have resumed work again. Your files will automatically be kept up-to-date. 
